---
title: "mini project_s18822"
output: html_document
date: "2024-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
library(skimr)
library(janitor)
library(zoo)
library(VIM)
library(imputeTS)
library(corrplot)
library(RColorBrewer)
library(gridExtra)
```

```{r}
london_air_data<-read_csv("../Data/london_local_data_2022.csv")
```
```{r}
london_local_sites<-read_csv("../Data/london_local_sites.csv")
```
```{r}
london_local_sites %>% filter(Parameter_name=="Ozone") %>% distinct(code)
london_local_sites %>% filter(Parameter_name=="Nitrogen dioxide") %>% distinct(code)
london_local_sites %>% filter(Parameter_name=="PM10 particulate matter (Hourly measured)") %>% distinct(code)


```
```{r}
london_air_data %>% distinct(code,site)
```
the data set has 34 sites
```{r}
sum(duplicated(london_air_data_new))
```
```{r}
london_air_data %>% filter(!is.na(no)) %>% distinct(code)
london_air_data %>% filter(!is.na(no2)) %>% distinct(code)
london_air_data %>% filter(!is.na(nox)) %>% distinct(code)
```





```{r}
london_air_data_new<-london_air_data %>% 
  mutate(dateNew=as.Date(date),time=format(as.POSIXct(date),format="%H:%M:%S"))
```

```{r}
merge_dataset<-merge(london_air_data_new,london_local_sites,by="code")
```



```{r}
glimpse(london_air_data_new)
```
```{r}
colSums(is.na(london_air_data_new))
```
```{r}
aggr(london_air_data_new)
```
```{r}
london_air_data_new1 <- london_air_data_new
```

```{r}
london_air_data_new1 <- lapply(london_air_data_new1, function(x) if (is.numeric(x)) na.approx(x) else x)
london_air_data_new1<-as.data.frame(london_air_data_new1)
#london_air_data_new1 <- london_air_data_new1 %>%
#  mutate_all(~ifelse(is.numeric(.), na.approx(.), .))

colSums(is.na(london_air_data_new1))

```
```{r}
str(london_air_data_new1)
```
```{r}
london_air_data_new1$date=as.POSIXct(london_air_data_new1$date)
london_air_data_new1$dateNew=as.POSIXct(london_air_data_new1$dateNew)
```



```{r}
london_air_data_new %>% filter(is.na(no2)) %>% distinct(code)
```
```{r}
london_air_data_new %>% group_by(date) %>% summarise(avg_hrly=mean(no2,na.rm=T))
```
plots
considering time
NO.NOx,No2

plot 1-

```{r}
data2<-london_air_data_new %>% 
  group_by(dateNew) %>% 
  summarise(avg_no=mean(no,na.rm=T),
            avg_nox=mean(nox,na.rm = T),
            avg_no2=mean(no2,na.rm=T))
```

```{r}
plot_no<-data2 %>% ggplot(aes(x=dateNew,y=avg_no))+geom_line(color="dark green")+labs(x="Date",y="average No concentration",subtitle = "average NO values throughout the year")+theme_minimal()
plot_no
```
```{r}
data2 %>% filter(avg_no==max(avg_no,na.rm = T)) %>% 
  select(dateNew)
```
```{r}
data2 %>% arrange(desc(avg_no)) %>%
  select(dateNew,avg_no) %>% head(10)
```

plot2:nOx
```{r}
plot_nox<-data2 %>% ggplot(aes(x=dateNew,y=avg_nox))+geom_line()+labs(x="Date",y="average NOx concentration",subtitle = "average NOx values throughout the year")+theme_minimal()
plot_nox
```
```{r}
data2 %>% filter(avg_nox==max(avg_nox,na.rm = T)) %>% 
  select(dateNew)
```

plot3:NO2
```{r}
plot_no2<-data2 %>% ggplot(aes(x=dateNew,y=avg_no2))+geom_line(color="dark green")+labs(x="Date",y="average NO2 concentration",subtitle = "average NO2 values throughout the year")+theme_minimal()
plot_no2
```
```{r}
data2 %>% filter(avg_no2==max(avg_no2,na.rm = T)) %>% 
  select(dateNew)
```
```{r}
data2 %>% arrange(desc(avg_no2)) %>%
  select(dateNew,avg_no2) %>% 
  head(10)
```


pm2-5 plot
```{r}
pm2_5_data<-london_air_data_new %>%group_by(code) %>% 
  filter(code =="HK6") %>% select(code,date,pm2_5,dateNew,time)
pm2_5_data
```
data imputaion for pm25 data in code HK6
```{r}
aggr(pm2_5_data)
sum(is.na(pm2_5_data$pm2_5))
```
```{r}
pm2_5_data %>% mutate(pm2_5=na_interpolation(pm2_5))
imputed<-na_interpolation(pm2_5_data$pm2_5)
sum(is.na(imputed))
pm2_5_data$pm2_5<-imputed
sum(is.na(pm2_5_data$pm2_5))
pm2_5_data %>% ggplot(aes(x=dateNew,y=pm2_5))+geom_line()+ggtitle("pm  2.5 concentration for code HK6")





```


since the pm2_5 data are measured only for 3 sites,extract data of those 3 sites 
```{r}
codes_to_filter<-c("HK6","HS4","WMD")
pm2_5_data_for_codes <- london_air_data_new %>%
  filter(code %in% codes_to_filter) %>%
  select(code, date, pm2_5, dateNew, time)
```

missing values in pm2_5 data
```{r}
colSums(is.na(pm2_5_data_for_codes))
```
interpolate missing values and plot 

```{r}
data3<-pm2_5_data_for_codes %>%  
  mutate(pm2_5_interpolate=na.approx(pm2_5))
```
#plot 4

```{r}
plot4_data<-data3 %>% group_by(dateNew) %>% 
  summarise(avg_pm2.5=mean(pm2_5_interpolate)) 
plot4_data%>% 
  ggplot(aes(x=dateNew,y=avg_pm2.5))+geom_line(color="dark green")+ geom_point(data=plot4_data %>%
               filter(avg_pm2.5 == max(avg_pm2.5, na.rm = TRUE)),
             aes(x = dateNew, y = avg_pm2.5),
             color = "red", size = 3) +
  labs(x="Date",y="average(daily) pm2.5 concentration",subtitle = "average(daily) PM2.5 values throughout the year")
```
```{r}
plot4_data %>% arrange(desc(avg_pm2.5)) %>% head(20)
```
2022-12-16 records the highest average pm2_5 concentration for the year which is slightly above the index

plot for so2
```{r}
london_air_data_new %>% filter(!is.na(so2)) %>% distinct(code)
```
so2 paramter is measured only at site LB5

```{r}
so2_data_for_codes <- london_air_data_new %>%
  filter(code == "LB5") %>%
  select(code, date, so2, dateNew, time) 
sum(is.na(so2_data_for_codes$so2))
```
there are 298 missing values, interpolate the missing values and plot

#plot5

```{r}
plot5_data<-so2_data_for_codes %>%  
  mutate(so2_interpolate=na.approx(so2)) %>% 
  group_by(dateNew) %>% 
  summarise(avg_daily_so2=mean(so2_interpolate)) 
plot5_data%>% 
  ggplot(aes(x=dateNew,y=avg_daily_so2))+geom_line()+labs(x="Date",y="average(daily) SO2 concentration",subtitle = "average(daily) SO2 values throughout the year")
```
```{r}
plot5_data %>% arrange(desc(avg_daily_so2)) %>% head(10)
```

plot for o3
```{r}
codes_to_filter<-c("HK6","R12","SK6")
o3_data_for_codes <- london_air_data_new %>%
  filter(code %in% codes_to_filter) %>%
  select(code, date, o3, dateNew, time)
sum(is.na(o3_data_for_codes$o3))
```
```{r}
plot6_data<-o3_data_for_codes %>%  
  mutate(o3_interpolate=na.approx(o3)) %>% 
  group_by(dateNew) %>% 
  summarise(avg_daily_o3=mean(o3_interpolate))
plot6_data%>% 
  ggplot(aes(x=dateNew,y=avg_daily_o3))+geom_line(color="dark green")+
  geom_point(data=plot6_data %>%
               filter(avg_daily_o3 == max(avg_daily_o3)),
             aes(x = dateNew, y = avg_daily_o3),
             color = "red", size = 3) +
  labs(x="Date",y="average(daily) O3 concentration",subtitle = "average(daily) O3 values throughout the year")
```
#spatial behaviour of pollutants
Top 2 polluted cities with respect to No2
```{r}
mean_by_code<-london_air_data_new %>% 
  group_by(code) %>% 
  summarise(avg_no2_by_code=mean(no2,na.rm = T),
            avg_no_by_code=mean(no,na.rm = T),
            avg_nox_by_code=mean(nox,na.rm = T))


```

#plot7:NO2 levels in sites

```{r}
plot7<-mean_by_code %>% 
  filter(!is.na(avg_no2_by_code)) %>%
  ggplot(aes(x = reorder(code,-avg_no2_by_code),y=avg_no2_by_code,fill=code))+geom_bar(stat="identity")+coord_flip()+labs(x="site(codes)",y="average NO2 concentration",subtitle = "NO2 levels in sites(increasing order)")+guides(fill=F)
plot7
```
#plot8:NO levels in sites

```{r}
plot8<-mean_by_code %>%
  filter(!is.na(avg_no_by_code)) %>% 
  ggplot(aes(x=reorder(code,-avg_no_by_code),y=avg_no_by_code,fill=code))+geom_bar(stat = "identity")+coord_flip()+theme_minimal()+labs(x="site(code)",y="average NO concentration",subtitle = "NO levels in sites(increasing order)")+guides(fill=F)
plot8
```
```{r}
grid.arrange(plot7, plot8, ncol=2)
```


```{r}
mean_by_code %>% ggplot(aes(x=reorder(code, -avg_nox_by_code),y=avg_nox_by_code))+geom_bar(stat = "identity")+coord_flip()+theme_minimal()
```
#plot9 :pm 2.5 levels for sites

```{r}
plot9<-london_air_data_new %>%
  group_by(code) %>% 
  summarise(mean_pm2_5=mean(pm2_5,na.rm=T)) %>% 
   filter(!is.na(mean_pm2_5)) %>% 
  ggplot(aes(x=code,y=mean_pm2_5,fill=code))+geom_bar(stat="identity")+theme_minimal()+labs(x="site(code)",y="average PM2.5 concentration",subtitle = "PM2.5 levels in sites(increasing order)")
plot9
```
```{r}
plot10<-london_air_data_new %>% 
  group_by(code) %>% 
 summarise(avg_o3_by_code=mean(o3,na.rm=T)) %>% 
   filter(!is.na(avg_o3_by_code)) %>% 
  ggplot(aes(x=code,y=avg_o3_by_code,fill=code))+geom_bar(stat="identity")+theme_minimal()+
  labs(x="site(code)",y="average O3 concentration",subtitle = "O3 levels in sites(increasing order)")
plot10
```
```{r}
pm10_data<-london_air_data_new %>% filter(!is.na(pm10)) %>% distinct(code)
#codes_to_filter_pm10<-as.vector(pm10_data)


pm10_data_for_codes <- london_air_data_new %>%
  filter(code %in% pm10_data) %>%
  select(code, date, pm10, dateNew, time)

pm10_data_for_codes %>%
  mutate(pm10_interpolate=na.approx(pm10)) %>% 
  group_by(code) %>% 
  summarise(avg_pm10_by_code=mean(pm10_interpolate)) %>% 
  ggplot(aes(x=code,y=avg_pm10_by_code))+geom_bar(stat="identity")
  
  
  
  
 
```

lets take a closer look at WA7
```{r}
wa7_data<-london_air_data_new %>% 
  group_by(code) %>% 
 filter(code=="WA7") %>% 
  select(code,date,dateNew,time,no2,pm10,pm2_5,no,nox)  
 
 aggr(wa7_data)
```
```{r}

```

```{r}
wa7_data_interpolated<-wa7_data %>%  
  mutate(pm10_interpolate=na.approx(pm10)) %>% 
  group_by(dateNew) %>% 
  summarise(avg_daily_pm10=mean(pm10_interpolate)) 
plot4<-wa7_data_interpolated%>% 
  ggplot(aes(x=dateNew,y=avg_daily_pm10))+geom_line()+
  geom_point(data=wa7_data_interpolated %>%
               filter(avg_daily_pm10 == max(avg_daily_pm10, na.rm = TRUE)),
             aes(x = dateNew, y = avg_daily_pm10),
             color = "red", size = 3) 
plot4


```
```{r}
wa7_data_interpolated %>% arrange(desc(avg_daily_pm10)) %>% head(10)

```
```{r}
wa7_data_interpolated %>%
  mutate(prev_value = lag(avg_daily_pm10),
         next_value = lead(avg_daily_pm10)) %>%
  filter( avg_daily_pm10 > next_value) %>%
  ggplot(aes(x = dateNew, y = avg_daily_pm10)) +
  geom_line() +
  geom_point(color = "red", size = 1) 
```
```{r}
max_peak_value <- wa7_data_interpolated %>%
  filter(avg_daily_pm10 == max(avg_daily_pm10, na.rm = TRUE)) %>%
  pull(avg_daily_pm10)

wa7_data_interpolated %>%
  ggplot(aes(x = dateNew, y = avg_daily_pm10)) +
  geom_line() +
  geom_point(data = wa7_data_interpolated %>% filter(avg_daily_pm10 == max_peak_value),
             aes(x = dateNew, y = avg_daily_pm10),
             color = "red", size = 3) 
```
corelation matrix
```{r}
correlation_matrix <- cor(london_air_data_new[, c("no", "no2", "nox", "so2","pm2_5","pm10")], use = "pairwise.complete.obs")

correlation_matrix[is.na(correlation_matrix)] <- 0
correlation_matrix

```
```{r}
color_palette <- colorRampPalette(c("#67001f", "#b2182b", "#d6604d", "#f4a582", "#fddbc7", "#f7f7f7", "#d1e5f0", "#92c5de", "#4393c3", "#2166ac", "#053061"))(18)


corrplot(correlation_matrix, 
         method = "color", 
         col = color_palette,
         type = "upper",  
         addCoef.col = "white",  
         tl.cex = 0.9,  
        
)
```

